TE_bilingual_vocabulary_model: Prepare the keepers_data for analysis

### 2021-06-11 03_prepare_data.Rmd
codes to prepare data

* If the child produced a greater number of words in English than French, then English is the dominant language.
* If the child produced a greater numnber of words in French than English, then French is the dominant language.
```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # define lang_dom_vocab
  mutate(lang_dom_vocab = case_when(total_words_eng > total_words_fre ~ "English",
                                 TRUE ~ "French")) %>%
  # make lang_dom_vocab a factor with 2 levels
  mutate(lang_dom_vocab = as.factor(lang_dom_vocab)) 
```

@@

* total number of words in the dominant language
* total number of words in the non-dominant language
* total number of singlets in the dominant language
* total number of singlets in the non-dominant language
* total number of singlets (i.e., across both languages)
```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  mutate(
    # define total number of words in the DOMINANT language
    total_words_dom = case_when(total_words_eng > total_words_fre ~ total_words_eng,
                            TRUE ~ total_words_fre),
    # define total number of words in the NON-DOMINANT language
    total_words_nondom = case_when(total_words_eng > total_words_fre ~ total_words_fre,
                               TRUE ~ total_words_eng),
    # define total number of singlets in the DOMINANT language
    total_singlet_dom = case_when(total_words_eng > total_words_fre ~ eng_unique_words,
                                  TRUE ~ fre_unique_words),
    # define total number of singlets in the NON-DOMINANT language
    total_singlet_nondom = case_when(total_words_eng > total_words_fre ~ fre_unique_words,
                                     TRUE ~ eng_unique_words),
    # compute total number of singlets across both languages
    singlet_vocab = total_singlet_dom + total_singlet_nondom)
```

@@

```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # calculate Vocabulary Balance balance_vocab
  mutate(balance_vocab = total_words_nondom/(total_words_dom+total_words_nondom)) %>%
  # calculate Input Balance balance_input
  mutate(
    # define input dominance (with reference to the lang_dom_vocab defined by vocab size)
    lang_dom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_eng,
                              TRUE ~ lang_exp_fre),
    lang_dom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_fre,
                                 TRUE ~ lang_exp_eng),
    # calculate the balance score
    balance_input = lang_dom_input/(lang_dom_input+lang_dom_input))
```

>>>>

* If the child produced a greater number of words in English than French, then English is the dominant language.
* If the child produced a greater numnber of words in French than English, then French is the dominant language.

```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # define lang_dom_vocab
  mutate(lang_dom_vocab = case_when(total_words_eng > total_words_fre ~ "English",
                                 TRUE ~ "French")) %>%
  # make lang_dom_vocab a factor with 2 levels
  mutate(lang_dom_vocab = as.factor(lang_dom_vocab)) 
```

@@

* total number of words in the dominant language
* total number of words in the non-dominant language
* total number of singlets in the dominant language
* total number of singlets in the non-dominant language
* total number of singlets (i.e., across both languages)

```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  mutate(
    # define total number of words in the DOMINANT language
    total_words_dom = case_when(total_words_eng > total_words_fre ~ total_words_eng,
                            TRUE ~ total_words_fre),
    # define total number of words in the NON-DOMINANT language
    total_words_nondom = case_when(total_words_eng > total_words_fre ~ total_words_fre,
                               TRUE ~ total_words_eng),
    # define total number of singlets in the DOMINANT language
    total_singlet_dom = case_when(total_words_eng > total_words_fre ~ eng_unique_words,
                                  TRUE ~ fre_unique_words),
    # define total number of singlets in the NON-DOMINANT language
    total_singlet_nondom = case_when(total_words_eng > total_words_fre ~ fre_unique_words,
                                     TRUE ~ eng_unique_words),
    # compute total number of singlets across both languages
    singlet_vocab = total_singlet_dom + total_singlet_nondom)
```

@@

```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # calculate Vocabulary Balance balance_vocab
  mutate(balance_vocab = total_words_nondom/(total_words_dom+total_words_nondom)) %>%
  # calculate Input Balance balance_input
  mutate(
    # define input dominance (with reference to the lang_dom_vocab defined by vocab size)
    lang_dom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_eng,
                              TRUE ~ lang_exp_fre),
    lang_nondom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_fre,
                                 TRUE ~ lang_exp_eng),
    # calculate the balance score
    balance_input = lang_nondom_input/(lang_dom_input+lang_nondom_input))
```

# Append percentile data to the keepers_data
Finally, we add the percentile data obtained from wordbankr to the keepers_data. This percentile data indicates the number of potentially learnable words for a child at the relevant age.

```{r}
# read in the percentile data
ws_percentile <- read.csv(here::here("data_keepers/ws_percentile.csv"))

# merge the percentile data to the keepers_ws_TE_final data
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # as the upper age limit of the CDI-WS is 30m, create a new age variable which replaces all >30m as 30m for merging with the percentile dataframe (this means the percentile should be the same for all >30m)
  mutate(age_months_percentile = ifelse(age_months_binned > 30, 30, age_months_binned)) %>%
  # merge with the percentile data frame
  left_join(ws_percentile, by = "age_months_percentile") 
```
